---
title: "mlr: Curve Data"
date: "`r Sys.Date()`"
output: html_document
---



# mlr: Curve Data

This is a short demo for the `mlr` branch `curve\_data`. The main goal of `curve\_data ` is to provide a tool for processing curvature data, i.e. data with temporal structure, within `mlr`. To create a non-temporal representation of the data, feature extraction methods are used. Currently, fourier feature transformation and discrete wavelet feature transformation are available within `curve_data`. The major structure of the `mlr` workflow remains unchanged. 
As example data, we use the gun-point data set available from the UCR time series archive (www.cs.ucr.edu/~eamonn/time_series_data/), which is a binary classification problem, with the given train-test split. 

```{r}
library(devtools)
library(BBmisc)
load_all()

## Load curve data
gp = load2("gunpoint.RData")
```


## Non temporal 

First, handling curvature data when ignoring the temporal structure, i.e. the temporal measurements are treated as (independent) covariates. The first step is to create a time series classification task in which the relevant information is stored, i.e. data, target variable, etc .

```{r}
## Define the task:
taskCl = makeTimeSeriesClassifTask(data = gp, target = "X1", positive = "1")
print(taskCl)

```

Next, the basic `mlr` step is to define which learner is used for classification. We choose a CART tree. The last steps are training the model i.e. the learner, predicting and evaluation. 

```{r}
## Define the learner:
lrn = makeLearner("classif.rpart")
print(lrn)


## Train the model on the training data of specified task:
model = train(lrn, taskCl, subset = 1:50)

## ...and predict for the test data:
pred = predict(model, taskCl, subset = 51:200)

## Get the mean misclassification error and true positive rate:
p = performance(pred, measures = list(mmce, tpr))
print(p)
```


## Wavelet transform

Now we use the `curve_data` functions to extract non-temporal features from the temporal data. To begin with, we choose the discrete wavelet feature transformation based on haar wavelets. 
For this, we first have to create a general time series classification task which is the input for the feature extraction method. We can specify which feature extraction method is used via the `method` parameter and add additional parameters (i.e. the `filter` and the `boundary`) in the `pars` argument. This functions returns a statndard classification task since the transformed data  does not inherit temporal structure anymore:

```{r}
## Define the general temporal task:
taskTs = makeTimeSeriesClassifTask(data = gp, target = "X1", positive = "1")

## Specify the feature extraction method and generate new task.
## Here, we use the Haar filter:
taskW = makeTSFeaturesClassifTask(task = taskTs, method = "wavelets", pars = list(filter ="haar"))
print(taskW)
```

We also want to evaluate a the Daubechie wavelet with filter length of 4. 

```{r}
## Use the Daubechie wavelet with filter length 4.
taskWD4 = makeTSFeaturesClassifTask(task = taskTs, method = "wavelets", pars = list(filter ="d4"))
print(taskWD4)
```

The following steps are equal to the non-temporal evaluation: train, predict and evaluate.

```{r}
## Train the model...:
model = train(lrn, taskW, subset = 1:50)

## ...and predict:
pred = predict(model, taskW, subset = 51:200)

## Get the mean misclassification error and true positive rate:
p = performance(pred, measures = list(mmce, tpr))
print(p)


## Do the same for the daubechie wavelet of length 4
model = train(lrn, taskWD4, subset = 1:50)
pred = predict(model, taskWD4, subset = 51:200)
p = performance(pred, measures = list(mmce, tpr))
print(p)
```


## Fourier transform

Now, we use the fourier feature transformation. Either the amplitude or the phase of the complex fourier coefficients can be used for analysis and can be specified in the additional `fft.coeff` argument:

```{r}
## Define the general temporal task:
taskTs = makeTimeSeriesClassifTask(data = gp, target = "X1", positive = "1")

## Specify the feature extraction method and generate  new task.
## We use the fourier features and the amplitude:
taskFa = makeTSFeaturesClassifTask(task = taskTs, method = "fourier", pars = list(fft.coeff = "amplitude"))
print(taskFa)

## ... or we prefer the phase:
taskFp = makeTSFeaturesClassifTask(task = taskTs, method = "fourier", pars = list(fft.coeff = "phase"))
print(taskFp)
```

The evaluation steps remain the same:

```{r}

## Train and predict. First, the amplitude features:
model = train(lrn, taskFa, subset = 1:50)
pred = predict(model, taskFa, subset = 51:200)
p = performance(pred, measures = list(mmce, tpr))
print(p)

## Train and predict with the phase features:
model = train(lrn, taskFp, subset = 1:50)
pred = predict(model, taskFp, subset = 51:200)
p = performance(pred, measures = list(mmce, tpr))
print(p)
```
