# Integrated Learners

In the following the learning methods already integrated into **mlr** are listed.

Columns **Num.**, **Fac.**, **NAs**, and **Weights** indicate if a method can cope with
numerical and factor predictors, if ``NA``s in the data are allowed and if observation
weights are supported.

Column **Props** shows further properties of the learning methods.
For *classification* you can see if binary and/or multi-class problems are supported.
In case of *survival analysis* the censoring type is shown. 
For example *rcens* means that the learning method can deal with right censored data.
Moreover, the type of prediction is displayed where *prob* indicates that probabilities
can be predicted.
In case of *regression* *se* means that standard errors and the mean response can be predicted.


```{r include=FALSE}
linkPkg = function(x) {
  collapse(sprintf("[%1$s](http://cran.r-project.org/web/packages/%1$s/)", x))
}

getTab = function(type) {
  lrns = listLearners(type, create = TRUE, warn.missing.packages = FALSE)
  cols = c("ID", "Short Name", "Name", "Packages", "Num.", "Fac.", "NAs", "Weights", "Props")
  df = makeDataFrame(nrow = length(lrns), ncol = length(cols),
    col.types = c("character", "character", "character", "character", "logical", "logical", "logical", "logical", "character"))
  names(df) = cols

  cn = function(x) if (is.null(x)) NA else x

  for (i in seq_along(lrns)) {
    lrn = lrns[[i]]
    df[i, 1] = lrn$id
    df[i, 2] = cn(lrn$short.name)
    df[i, 3] = cn(lrn$name)
    df[i, 4] = linkPkg(lrn$package)
    df[i, 5] = hasProperties(lrn, "numerics")
    df[i, 6] = hasProperties(lrn, "factors")
    df[i, 7] = hasProperties(lrn, "missings")
    df[i, 8] = hasProperties(lrn, "weights")
    df[i, 9] = collapse(setdiff(lrn$properties, c("numerics", "factors", "missings", "weights")))
  }
  logicals = vlapply(df, is.logical)
  df[logicals] = lapply(df[logicals], function(x) ifelse(x, "X", ""))
  pandoc.table(df, style = "rmarkdown", split.tables = Inf, split.cells = Inf,
    justify = c("left", "left", "left", "left", "center", "center", "center", "center", "right"))
}
```

### Classification
```{r echo=FALSE,results="asis"}
getTab("classif")
```

### Regression
```{r echo=FALSE,results="asis"}
getTab("regr")
```

### Survival analysis
```{r echo=FALSE,results="asis"}
getTab("surv")
```

### Cluster analysis
```{r echo=FALSE,results="asis"}
getTab("cluster")
```
