# Learning Curve Analysis

To analyse how the increase of observations in the training set improves the performance of
a learner the *learning curve* is an appropriate visual tool.
The experiment is conducted with an increasing subsample size and the performance is measured.
In the plot the x-axis represents the relative subsample size whereas the y-axis represents
the performance.

Note that this function internally uses [&benchmark] in combination with [&makeDownsampleWrapper],
so for every run new observations are drawn.
Thus the results are noisy.
To reduce noise increase the number of resampling iterations.
You can define the resampling method in the `resampling` argument of [&generateLearningCurveData].
It is also possible to pass a [ResampleInstance](&makeResampleInstance) (which is a result
of [&makeResampleInstance]) to make resampling consistent for all passed learners and each
step of increasing the number of observations.


## Plotting the learning curve

The [%mlr] function [&generateLearningCurveData] can generate the data for *learning curves*
for multiple [learners](integrated_learners.md) and multiple [performance measures](measures.md)
at once.
With [&plotLearningCurve] the result of [&generateLearningCurveData] can be plotted using [%ggplot2].
[&plotLearningCurve] has an argument `facet` which can be either "measure" or "learner". By default
`facet = "measure"` and facetted subplots are created for each measure input to [&generateLearningCurveData].
If `facet = "measure"` learners are mapped to color, and vice versa.

```{r LearningCurveTPFP}
r = generateLearningCurveData(
  learners = list("classif.rpart", "classif.knn"),
  task = sonar.task,
  percs = seq(0.1, 1, by = 0.2),
  measures = list(tp, fp, tn, fn),
  resampling = makeResampleDesc(method = "CV", iters = 5),
  show.info = FALSE)
plotLearningCurve(r)
```

What happens in [&generateLearningCurveData] is the following:
Each learner will be internally wrapped in a [DownsampleWrapper](&makeDownsampleWrapper).
To measure the performance at the first step of `percs`, say `0.1`, first the data will be
split into a *training* and a *test set* according to the given *resampling strategy*.
Then a random sample containing 10% of the observations of the *training set* will be drawn
and used to train the learner.
The performance will be measured on the *complete test set*.
These steps will be repeated as defined by the given *resampling method* and for each value
of `percs`.

In the first example a simplified usage of the `learners` argument was used, so that it's
sufficient to give the *name*.
It is also possible to create a learner the usual way and even to mix it.
Make sure to give different `id`s in this case.

```{r LearningCurveACCx, fig.width = 8, fig.height = 6}
lrns = list(
  makeLearner(cl = "classif.ksvm", id = "ksvm1" , sigma = 0.2, C = 2),
  makeLearner(cl = "classif.ksvm", id = "ksvm2" , sigma = 0.1, C = 1),
  "classif.randomForest"
)
rin = makeResampleDesc(method = "CV", iters = 5)
lc = generateLearningCurveData(learners = lrns, task = sonar.task,
                               percs = seq(0.1, 1, by = 0.1), measures = acc,
                               resampling = rin, show.info = FALSE)
plotLearningCurve(lc)
```

There is also an experimental [%ggvis] plotting function, [&plotLearningCurveGGVIS]. Instead of the `facet`
argument to [&plotLearningCurve] there is an argument `interactive` which plays a similar role. As subplots
are not available in [%ggvis], measures or learners are mapped to an interactive sidebar which allows selection
of the displayed measures or learners. The other feature is mapped to color.

```{r, eval = FALSE}
plotLearningCurveGGVIS(r, interactive = "measure")
```
