Learners
========

These classes provide a unified interface to all machine learning methods in R
-- classification, regression, survival analysis, and clustering.
Many are already integrated,
others are not, but the package is specifically designed to make
extensions simple. If your favorite methods is missing, either
send the **mlr** maintainers an email or see the part of the tutorial
on [how to extend the package yourself](create_learner.md).

To see which learners are already implemented, have a look at the \man[learners] documentation page.

A learner in **mlr** is constructed by calling \man[makeLearner].
In the constructor you can do the following things:

* Set hyperparameters.
* Set the ID to name the object (some methods will later use this ID to name results).
* Control the output for later prediction, e.g. for classification
  whether you want a factor for predicted classes or probabilities.

```{r}
# classifier, set it up for prediction probabilities
lrn = makeLearner("classif.rpart", predict.type = "prob")
# regression
lrn = makeLearner("regr.rpart")
# gradient boosting machine with hyperparameters and custon name
lrn = makeLearner("regr.gbm", id = "mygbm", n.trees = 500, interaction.depth = 3)
```

The first argument specifies which algorithm to use.
The naming convention is ``classif.<R_method_name>`` for
classification methods, ``regr.<R_method_name>`` for regression methods,
``surv.<R_method_name>`` for survival analysis, and ``cluster.<R_method_name>``
for clustering methods.
The tables in [IntegratedLearners](integrated_learners.md) provide more
information on the properties for all integrated learning methods.

Further information
-------------------

The generated learner is an object of class \man2[Learner][makeLearner].
It contains the properties the method, e.g. which types of features it can handle,
what can be output during prediction, whether multi-class problems,
observations weights or missing values are supported and so on.

```{r}
# classifier, set it up for prediction probabilities
lrn = makeLearner("classif.rpart", predict.type = "prob", minsplit = 20)
lrn
# display stored information for hyperparameters
lrn$par.set
```

