# Multi-criteria Evaluation

In this example, we have to deal with some of the implementation details of
**mlr** and the characteristics of different measures.
In a lot of cases you get more than just one measure value but are interested in
just one aggregated value.
For example, a 10-fold cross validation computes 10 values for the chosen
performance measure.
The aggregated value is the mean of these 10 numbers.
**mlr** knows how to handle it because each `measure` knows how it is aggregated:

```{r MulticriteriaEvaluation_aggr}
library(mlr)
mmce$aggr #Mean misclassification error
rmse$aggr #Root mean square error
```

You can also create a new `measure` by using a different aggregation (see \man[aggregations]).

```{r MulticriteriaEvaluation_NewAggr}
mmceAggrBySd = setAggregation(mmce, test.sd)

task = makeClassifTask(data = iris, target = "Species")
lrn = makeLearner("classif.rpart")
rdesc = makeResampleDesc("CV", iters = 5)
r = resample(lrn, task, rdesc, measures = list(mmce, mmceAggrBySd))
```

It is even possible to create your own aggregation function by calling the
internal **mlr** function \man[makeAggregation].
You can use internal (not exported) functions of R packages by prefixing the
name of the function with the name of the package, i.e. `packagename:::function()`.
In this case:
```{r MulticriteriaEvaluation_makeAggregation}
mlr:::makeAggregation(id = "some.name",
  fun = function (task, perf.test, perf.train, measure, group, pred) {
    #stuff you want to do with perf.test or perf.train
  })
```
Remember: it is important that the head of the function looks exactly as above!
`perf.test` and `perf.train` are both numerical vectors containing the measure values.
In the usual case (e.g. cross validation), the `perf.train` vector is empty.

## Practical Example: Evaluate the Range of Measures

Let's say you are interested in the range of the obtained measures:
```{r MulticriteriaEvaluation_range}
my.range.aggr = mlr:::makeAggregation(id="test.range",
  fun = function (task, perf.test, perf.train, measure, group, pred) diff(range(perf.test)))
```

Now we can run a feature selection based on the first measure in the provided
list and see how the other measures turn out.
```{r MulticriteriaEvaluation}
library(ggplot2)
ms1 = mmce
ms2 = setAggregation(ms1, my.range.aggr)
ms1min = setAggregation(ms1, test.min)
ms1max = setAggregation(ms1, test.max)
(res = selectFeatures(lrn, task, rdesc, measures=list(ms1, ms2, ms1min, ms1max),  control=makeFeatSelControlExhaustive(), show.info=FALSE))
perf.data = as.data.frame(res$opt.path)
p = ggplot(aes(x=mmce.test.mean, y=mmce.test.range, xmax=mmce.test.max, xmin=mmce.test.min, color=as.factor(Sepal.Width), pch=as.factor(Petal.Width)), data=perf.data) +
  geom_point(size=4) +
  geom_errorbarh(height=0)
print(p)
```
