# Parallelization

**R** by default does not make use of parallelization.
With the integration of [%parallelMap] into [%mlr], it becomes easy to activate the parallel computing capabilities already supported by [%mlr].
[%parallelMap] supports all major parallelization backends: local multicore execution using [%parallel], socket and MPI clusters using [%snow], makeshift SSH-cluster using [%BatchJobs] and high performance computing clusters (managed by a scheduler like SLURM, Torque/PBS, SGE or LSF) also using [%BatchJobs].

All you have to do is select a backend by calling one of the
[parallelStart*](&parallelMap::parallelStart) functions.
The first loop [%mlr] encounters which is marked as parallel executable will be automatically parallelized.
It is good practice to call [parallelStop](&parallelMap::parallelStop) at the end of your script.

```{r}
library("parallelMap")
parallelStartSocket(2)
rdesc = makeResampleDesc("CV", iters = 3)
r = resample("classif.lda", iris.task, rdesc)
parallelStop()
```

On Linux or Mac OS X, you may want to use
[parallelStartMulticore](&parallelMap::parallelStart) instead.
We offer different parallelization levels for fine grained control over the parallelization.
E.g., if you do not want to parallelize the [&benchmark] function because it has only very few iterations but want to parallelize the [resampling](&resample) of each learner instead,
you can specifically pass the ``level`` `"mlr.resample"` to the [parallelStart*](&parallelMap::parallelStart) function.
Currently the following levels are supported:
```{r}
parallelGetRegisteredLevels()
```

Here is a brief explanation of what these levels do:

* **resample**: Each resampling iteration (a train / test step) is a parallel job.
* **benchmark**: Each experiment "run this learner on this data set" is a parallel job.
* **tuneParams**: Each evaluation in hyperparameter space "resample with these parameter settings" is a parallel job.
  How many of these can be run independently in parallel, depends on the tuning algorithm.
  For grid search or random search this is
  no problem, but for other tuners it depends on how many points are produced in each iteration
  of the optimization. If a tuner works in a purely sequential fashion, we cannot work magic and the
  hyperparameter evaluation will also run sequentially. But note that you can still parallelize the
  underlying resampling.
* **selectFeatures**: Each evaluation in feature space "resample with this feature subset" is a parallel job.
  The same comments as for "tuneParams" apply here.

For further details, consult the
[parallelMap tutorial](https://github.com/berndbischl/parallelMap#parallelmap) and help.
