# Parallelization

*R* by default doesn't make use of multiple cores.
With the integration of [parallelMap](https://github.com/berndbischl/parallelMap) into **mlr**, it becomes easy to activate the parallel computing capabilities already supported by **mlr**.

All you have to do is select an interface by calling one of the [parallelStart*](&parallelMap::parallelStart) functions.
The first annotated loop **mlr** encounters will then be automatically parallelized.
Afterwards it is good practice to call [&parallelMap::parallelStop].

```{r}
library("parallelMap")
parallelStartSocket(2)
task = makeClassifTask(data = iris, target = "Species")
lrn = makeLearner("classif.lda")
rdesc = makeResampleDesc("CV", iters = 4)
r = resample(lrn, task, rdesc)
parallelStop()
```

In Linux or Mac OS X, you may want to use `parallelStartMulticore()` instead.
We over different parallelization levels for fine grain control over the parallelization.
E.g., if you do not want to parallelize the `benchmark()` function because it has only very few iterations but want to parallelize the resampling of each learner instead, you can specifically pass the level `"resample"` to the `parallelStart*` function.
Currently the following levels are supported:
```{r}
parallelShowRegisteredLevels()
```

For further ways of parallelizing your computations, consult the [parallelMap tutorial](https://github.com/berndbischl/parallelMap#parallelmap) and help.
