# Data Preprocessing

[%mlr] offers several options for data preprocessing.

Some of the following simple methods were already mentioned in section
[Learning tasks](task.md):

* [&capLargeValues]: Convert large/infinite numeric values in a [data.frame](&base::data.frame)
  or [&Task].
* [&createDummyFeatures]: Generate dummy variables for factor features in a
  [data.frame](&base::data.frame) or [&Task].
* [&dropFeatures]: Remove some features from a [&Task].
* [&joinClassLevels]: Only for classification: Merge existing classes in a [&Task] to
  new, larger classes.
* [&mergeSmallFactorLevels]: Merge infrequent levels of factor features in a [&Task].
* [&normalizeFeatures]: Normalize features in a [&Task] by different methods, e.g.,
  standardization or scaling to a certain range.
* [&removeConstantFeatures]: Remove constant features from a [&Task].
* [&subsetTask]: Remove observations and/or features from a [&Task].

Moreover, distinct sections in this tutorial are devoted to

* [Feature selection](feature_selection.md) and
* [Imputation of missing values](impute.md).

Additionally, [%mlr] permits to fuse a [Learner](&makeLearner) with any preprocessing method
of your choice like any kind of data transformation, normalization, dimensionality reduction
or outlier removal.


## Fusing a learner with data preprocessing

A [Learner](&makeLearner) can be coupled with a preprocessing method by function 
[&makePreprocWrapper].

As described in the section about [wrapped learners](wrapper.md) wrappers are implemented
using a *train* and a *predict* method.
In case of preprocessing wrappers these methods specify how to transform the data before
training and prediction and are *completely defined by the user*.

The specified preprocessing steps then "belong" to the wrapped [Learner](&makeLearner).
In contrast to the preprocessing options listed above like [&normalizeFeatures]

* the [&Task] itself remains unchanged,
* the preprocessing is not done globally, i.e., for the whole data set, but for every pair of
  training/test data sets in e.g. resampling,
* any parameters controlling the preprocessing as, e.g., the percentage of outliers to be removed
  can be [tuned](tune.md) together with the base learner parameters.

Let's see how to create a preprocessing wrapper using the following simple example:
Some learning methods as e.g. k nearest neighbors, support vector machines or neural networks
usually require scaled features.
Many, but not all, have a built-in scaling option where the training data set is scaled before
model fitting and the test data set is scaled accordingly, that is by using the scaling
parameters from the training stage, before making predictions.
In the following we show how to add a scaling option to a [Learner](&makeLearner) by coupling
it with function [scale](&base::scale).


### Specifying the train method

The *train* function has to be a function with the following arguments:

* ``data`` is a [data.frame](&base::data.frame) with values of all features and
  the target variable.
* ``target`` is a string and denotes the name of the target variable in ``data``.
* ``args`` is a [list](&base::list) of further arguments and parameters that influence the
  preprocessing.

It must return a [list](&base::list) with elements ``$data`` and ``$control``,
where ``$data`` is the preprocessed data set and ``$control`` stores all information required
to preprocess the data before prediction.

The *train* function for the scaling example is given below. It calls [scale](&base::scale) on the
numerical features and returns the scaled training data and the corresponding scaling parameters.
``args`` contains the ``center`` and ``scale`` arguments of function [scale](&base::scale)
and slot ``$control`` stores the scaling parameters.


```{r}
tr.fun = function(data, target, args = list(center, scale)) {
  cns = colnames(data)
  nums = setdiff(cns[sapply(data, is.numeric)], target)
  x = as.matrix(data[, nums, drop = FALSE])
  x = scale(x, center = args$center, scale = args$scale)
  ctrl = args
  if (is.logical(ctrl$center) && ctrl$center)
    ctrl$center = attr(x, "scaled:center")
  if (is.logical(ctrl$scale) && ctrl$scale)
    ctrl$scale = attr(x, "scaled:scale")
  data = data[, setdiff(cns, nums), drop = FALSE]
  data = cbind(data, as.data.frame(x))
  return(list(data = data, control = ctrl))
}
```


### Specifying the predict method

The *predict* function has the following arguments:

* ``data`` is a [data.frame](&base::data.frame) with feature values.
  (Naturally, it does not contain values of the target variable.)
* ``target`` is a string indicating the name of the target variable.
* ``args`` are the ``args`` that were passed to the *train* function.
* ``control`` is the object returned by the *train* function.

It returns the preprocessed data.

In our running example the *predict* function scales the numerical features using the
parameters from the training stage stored in ``control``.

```{r}
pr.fun = function(data, target, args, control) {
  cns = colnames(data)
  nums = cns[sapply(data, is.numeric)]
  x = as.matrix(data[, nums, drop = FALSE])
  x = scale(x, center = control$center, scale = control$scale)
  data = data[, setdiff(cns, nums), drop = FALSE]  
  data = cbind(data, as.data.frame(x))
  return(data)
}
```


### Creating the preprocessing wrapper

Below we create a preprocessing wrapper with a [regression neural network](&nnet::nnet) (which
itself does not have a scaling option) as base learner.

The *train* and *predict* functions defined above are passed to [&makePreprocWrapper] via
the ``train`` and ``predict`` arguments.
``par.vals`` is a [list](&base::list) of parameter values that is relayed to the ``args``
argument of the *train* function.

```{r}
lrn = makeLearner("regr.nnet", trace = FALSE, decay = 1e-02)
lrn = makePreprocWrapper(lrn, train = tr.fun, predict = pr.fun,
  par.vals = list(center = TRUE, scale = TRUE))
lrn
```

Let's compare the cross-validated mean squared error ([mse](&measures)) on the
[Boston Housing data set](&mlbench::BostonHousing) with and without scaling.

```{r}
rdesc = makeResampleDesc("CV", iters = 10)

r = resample(lrn, bh.task, resampling = rdesc, show.info = FALSE)
r$aggr

lrn = makeLearner("regr.nnet", trace = FALSE, decay = 1e-02)
r = resample(lrn, bh.task, resampling = rdesc, show.info = FALSE)
r$aggr
```{r}
