# Learning Tasks

Learning tasks are the basic elements of the package that encapsulate the
data set and all relevant information regarding the purpose of the
task, e.g., the target variable.

The tasks are organised in a hierarchy, with the generic
[&Task] at the top.
The concrete tasks that can be instantiated are [ClassifTask](&Task) for classification,
[RegrTask](&Task) for regression, [SurvTask](&Task) for survival
analysis, [CostSensTask](&Task) for
cost-sensitive classification, and [ClusterTask](&Task) for clustering.


## Creating a learning task

There are dedicated functions to create these tasks:
[makeClassifTask](&Task),
[makeRegrTask](&Task),
[makeSurvTask](&Task), [makeCostSensTask](&Task), and
[makeClusterTask](&Task).
These require the data set and, depending on the 
nature of the learning problem, further information.


### Cluster analysis

In order to generate a task for a cluster analysis problem ``data`` is the only mandatory argument.
In the following example we create a learning task from the data set [mtcars](&datasets::mtcars).

```{r}
cluster.task = makeClusterTask(id = "cars", data = mtcars)
cluster.task
```

The argument ``id`` is optional and allows to specify a custom name for the task.
It will be later used to name results, for example of
[benchmark experiments](benchmark_experiments.md).
If not specified the ``id`` defaults to the name of the **R** object passed to ``data``,
in the above example ``"mtcars"``.


### Regression

For supervised learning like regression (as well as classification and
survival analysis) we, in addition to ``data``, have to specify the name of the ``target``
variable.

```{r}
data(BostonHousing, package = "mlbench")

regr.task = makeRegrTask(data = BostonHousing, target = "medv")
regr.task
```

As you can see, the task records the type of the learning problem and basic information about
the data set, e.g., the types of the features (numeric vectors or factors),
the number of observations, or whether missing values are present.

Creating tasks for (cost-sensitive) classification problems and survival analysis
follows the same scheme, except that the target variable is a factor instead of a numeric quantity of course.
For each of these learning problems, some specifics are described below.


### Classification

For classification the target variable has to be a ``factor``.

In binary classification it is customary to refer to the two classes as *positive* and
*negative* class.
This is for example relevant with regard to certain [performance measures](performance.md)
like the *true positive rate*.
By default the first factor level of the target variable is selected as the positive class.

In the following example, we define a classification task for the data set 
[BreastCancer](&mlbench::BreastCancer) and exclude the variable ``Id`` from all further
model fitting and evaluation.

```{r}
data(BreastCancer, package = "mlbench")

df = BreastCancer
df$Id = NULL
classif.task = makeClassifTask(id = "BreastCancer", data = df, target = "Class")
classif.task
```

The positive class is benign.
Class malignant can be manually selected as the positive class:

```{r}
classif.task = makeClassifTask(id = "BreastCancer", data = df, target = "Class",
    positive = "malignant")
```


### Survival analysis

For survival analysis ``target`` specifies the names of the survival time and event
columns in ``data`` and therefore is a character vector of length 2.
Note that the event column has to be a logical vector.
For details see the documentation of function [Surv](&survival::Surv).

```{r}
data(lung, package = "survival")
## Coerce the event column to a logical vector
lung$status = lung$status - 1

surv.task = makeSurvTask(data = lung, target = c("time", "status"))
surv.task
```

The type of censoring can be specified via the argument ``censoring`` which defaults to
``"rcens"`` for right censored data.


### Cost-sensitive classification

In regular classification, as addressed above, the aim when training a classifier is to 
minimize the misclassification rate.
Thus, it is implicitly assumed that all types of misclassification errors are equally severe.
In *cost-sensitive classification* the costs of misclassifying an object with true
class *y* as class *k* can be different depending on *y* and *k*.

A further generalization is *cost-sensitive learning with example-specific costs*.
In this scenario, the misclassification costs do not only depend on the true and predicted
class labels, but also on the feature values *x*.
Thus, for a *K*-class problem, each example (*x*, *y*) is associated with a vector
of costs for predicting classes 1,...,*K* given the true class label *y*.
If the cost vectors are given, the true values *y* of the target variable not needed anymore.
Naturally, it is assumed that the cost of the true class label *y* is minimal.

Hence, in order to create a cost-sensitive classification task, the feature values *x* and
a ``cost`` matrix that contains the cost vectors for all examples in the data set are required. 


```{r}
## We use the iris data and generate an artificial cost matrix
## (following Beygelzimer et al. (2005))
df = iris
cost = matrix(runif(150 * 3, 0, 2000), 150) * (1 - diag(3))[df$Species,]
df$Species = NULL

costsens.task = makeCostSensTask(data = df, cost = cost)
costsens.task
```

Beygelzimer, A., Dani, V., Hayes, T., Langford, J., and Zadrozny, B.
Error limiting reductions between classification tasks.
In Machine Learning: Proceedings of the 22rd International Conference, pp. 49–56. ACM, 2005.

For more details see the section about [cost-sensitive classification](cost_sensitive_classif.md).


### Further details

The [&Task] help page also lists several other arguments to describe further details of the
problem.

For example, we could include a ``blocking`` factor into the task.
This would tell the task that some observations "belong together", and should
not be separated when splitting into training and test sets during a resampling iteration.

Another possibility is to ``weight`` observations according to their importance.


## Accessing a learning task

[&Task] is a ``list`` and its main properties are stored in element
``"task.desc"`` which is of class [&TaskDesc].

```{r}
str(regr.task$task.desc)
```

Moreover, there are some functions, following the naming convention
``getTask<what_to_extract>``, to access the data set and some of its
properties.
The most important ones are listed in the documentation of [&Task].
Here are some examples.

```{r}
## Get the number of input variables in cluster.task:
getTaskNFeats(cluster.task)

## Get the names of the input variables in cluster.task:
getTaskFeatureNames(cluster.task)

## Get the values of the target variable in regr.task:
head(getTaskTargets(regr.task))

## Get the cost matrix in costsens.task:
head(getTaskCosts(costsens.task))

## Accessing the data set in classif.task:
str(getTaskData(classif.task))
```

Note the many options [&getTaskData] provides to convert the data set into a convenient format.
This is especially handy when you [integrate a learner](create_learner.md) from another **R**
package into [%mlr].
In this regard the functions [&getTaskFormula] and [&getTaskFormulaAsString] are also useful.


## Modifying a learning task

[%mlr] provides several functions to alter an existing [&Task] which is often more
convenient than creating a new [&Task] from scratch.
Here are some examples.

```{r}
## Select observations and features:
cluster.task = subsetTask(cluster.task, subset = 4:17)

## It may happen, especially after selecting observations, that features are constant.
## These should be removed.
removeConstantFeatures(cluster.task)

## Remove selected features:
dropFeatures(surv.task, c("meal.cal", "wt.loss"))

## Standardize numerical features:
normalizeFeatures(classif.task, method = "standardize")
```


## Pre-defined tasks

For your convenience [%mlr] provides pre-defined [&Task]s for each type of learning problem.
These are used throughout this tutorial in order to get shorter and more readable code.

```{r echo=FALSE, results="asis"}
linkTask = function(x) {
    collapse(sprintf("[%1$s](http://www.rdocumentation.org/packages/mlr/functions/%1$s/)", x), sep = "<br />")
}
d = data(package = "mlr")
d = d$results
df = as.data.frame(d[seq(2,nrow(d),2), c("Item", "Title")])
df$Item = sapply(df$Item, linkTask)
names(df)[1] = "Task"
pandoc.table(df, style = "rmarkdown", split.tables = Inf, split.cells = Inf,
	justify = c("left", "left"))
```
