Training a Learner
==================

Training a learner means fitting a model to a given data set.
In the **mlr** package this can be done by calling the function \man[train]
with a learner and a suitable \man[Task].


Quick start
-----------

### Classification example

We start with a basic example and train a Linear Discriminant Analysis on the ``iris`` data set.

```{r}
library("mlr")
task = makeClassifTask(data = iris, target = "Species")
lrn = makeLearner("classif.lda")
mod = train(lrn, task)
mod
```

### Regression example

We fit a simple linear regression model to the ``BostonHousing`` data set.

```{r}
library("mlr")
library("mlbench")
data(BostonHousing)

task = makeRegrTask(data = BostonHousing, target = "medv")
lrn = makeLearner("regr.lm")
mod = train(lrn, task)
mod
```

### Clustering example

We cluster the ``iris`` data set without the target variable.

```{r}
library("mlr")
task = makeClusterTask(data = iris[,-5])
lrn = makeLearner("cluster.XMeans")
mod = train(lrn, task)
mod
```

Further information
-------------------

As already mentioned, we can train a learner simply by
calling the function \man[train], passing it a \man[Learner] and a \man[Task]. 
In the examples above, the machine learning method was specified 
through special \man2[Learner][makeLearner] objects. Creating a separate object
allows to set the hyperparameters of the learner before training for example.

Optionally, only a subset of the data, specified by an index set, can be used to 
train the learner. This set is passed using the ``subset`` argument of \man[train].

The return value is always an object of class \man[WrappedModel], which wraps the
particular model of the used R classification or regression method. It
can subsequently be used to perform a \man2[prediction][predict.WrappedModel] for new observations.
