---
title: "Use case gstat"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message=FALSE}
library("mlr")
library("sp")
library("gstat")
library("ParamHelpers")
library("dplyr")

## show grouped code output instead of single lines
knitr::opts_chunk$set(collapse = FALSE)
knitr::knit_hooks$set(document = function(x){
  gsub("```\n*```r*\n*", "", x)
})
```

From gstat CRAN [vignette](https://cran.r-project.org/web/packages/gstat/vignettes/gstat.pdf), let's compare raw-gstat outputs with mlr-gstat outputs.

## Prepare the required libraries
```{r}
library(gstat)
library(sp)
library(mlr)
library(dplyr)
```

##trend surfaces
### trend surfaces - raw gstat
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# making spatial
coordinates(meuse) = ~x+y
coordinates(meuse.grid) = ~x+y
gridded(meuse.grid) = TRUE
# interpolating
ts.raw <- meuse.grid
ts.raw$ts1 = krige(log(zinc) ~ 1, meuse, meuse.grid, degree = 1)$var1.pred
ts.raw$ts2 = krige(log(zinc) ~ 1, meuse, meuse.grid, degree = 2)$var1.pred
ts.raw$ts3 = krige(log(zinc) ~ 1, meuse, meuse.grid, degree = 3)$var1.pred
# mapping
ts.raw.plot <- spplot(ts.raw, c("ts1", "ts2", "ts3"), main = "log(zinc), trend surface interpolation")
ts.raw.plot
```

### trend surfaces - mlr gstat
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# adding a column with log zinc
meuse <- meuse %>% dplyr::mutate(log_zinc = log(zinc))
# defining the regression task
task = makeRegrTask(id = "meuse",  data = meuse, target = "log_zinc")
task.ts<- dropFeatures(task = task, features = getTaskFeatureNames(task)[-c(1,2)])
# defining the learner
lrn.ts.1 = makeLearner(cl = "regr.gstat", id= "mlr-ts1", degree = 1, locations = ~x+y)
lrn.ts.2 = makeLearner(cl = "regr.gstat", id= "mlr-ts2", degree = 2, locations = ~x+y)
lrn.ts.3 = makeLearner(cl = "regr.gstat", id= "mlr-ts3", degree = 3, locations = ~x+y)
# training the learners
mod.ts.1 = train(lrn.ts.1, task.ts)
mod.ts.2 = train(lrn.ts.2, task.ts)
mod.ts.3 = train(lrn.ts.3, task.ts)
# interpolating
ts.mlr <- meuse.grid
newdata.pred.ts.1 = predict(mod.ts.1, newdata = meuse.grid)
ts.mlr$mlr.ts.1<- (bind_cols(data.frame(meuse.grid), newdata.pred.ts.1$data))$response
newdata.pred.ts.2 = predict(mod.ts.2, newdata = meuse.grid)
ts.mlr$mlr.ts.2 <- (bind_cols(data.frame(meuse.grid), newdata.pred.ts.2$data))$response
newdata.pred.ts.3 = predict(mod.ts.3, newdata = meuse.grid)
ts.mlr$mlr.ts.3 <- (bind_cols(data.frame(meuse.grid), newdata.pred.ts.3$data))$response
# mapping
coordinates(ts.mlr) <- ~x+y
gridded(ts.mlr) = TRUE
ts.mlr.plot <- spplot(ts.mlr, c("mlr.ts.1", "mlr.ts.2", "mlr.ts.3"), main = "log(zinc), trend surface interpolation")
ts.mlr.plot
```
### identical predictions ? 
```{r}
identical(ts.mlr$mlr.ts.1, ts.raw$ts1)
identical(ts.mlr$mlr.ts.2, ts.raw$ts2)
identical(ts.mlr$mlr.ts.3, ts.raw$ts3)
```


## idw
### idw - raw gstat
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# making spatial
coordinates(meuse) = ~x+y
coordinates(meuse.grid) = ~x+y
gridded(meuse.grid) = TRUE
# interpolating
zinc.idw = idw(zinc~1, meuse, meuse.grid)
# mapping
idw.raw.plot <- spplot(zinc.idw["var1.pred"], do.log = F, colorkey=TRUE,  main = "zinc inverse distance weighted interpolations")
idw.raw.plot
```

### idw - mlr gstat
mlr only works with pure dataframes. neither tibbles, sp, or sf dataframes are supported. 
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# defining the regression task
task = makeRegrTask(id = "meuse",  data = meuse, target = "zinc")
task.idw <- dropFeatures(task = task, features = getTaskFeatureNames(task)[-c(1,2)])
# defining the learner
lrn.idw = makeLearner(cl = "regr.gstat", id= "mlr-idw", locations = ~x+y)
# training the model
mod.idw = train(lrn.idw, task.idw)
# interpolating
newdata.pred.idw = predict(mod.idw, newdata = meuse.grid)
mlr.idw <- bind_cols(data.frame(meuse.grid), newdata.pred.idw$data)
# mapping
coordinates(mlr.idw) <- ~x+y
gridded(mlr.idw) = TRUE
idw.mlr.plot <- spplot(mlr.idw["response"], do.log = F, colorkey = TRUE, main = mod.idw$learner$id)
idw.mlr.plot
```

### identical predictions ? 
```{r}
identical(zinc.idw["var1.pred"]@data[[1]], mlr.idw["response"]@data[[1]])
```

## ordinary kriging example
### ordinary kriging - raw gstat
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# making spatial
coordinates(meuse) = ~x+y
coordinates(meuse.grid) = ~x+y
gridded(meuse.grid) = TRUE
# computing sample variogram
lzn.vgm = variogram(log(zinc)~1, meuse)
# manually fitting a model to the vgm with constant mean
lzn.fit = fit.variogram(lzn.vgm, model = vgm(1, "Sph", 900, 1))
plot(lzn.vgm, lzn.fit)
# kriging
lzn.kriged = krige(log(zinc)~1, meuse, meuse.grid, model = lzn.fit)
# mapping
lzn.kriged.plot <- spplot(lzn.kriged["var1.pred"], do.log = F, colorkey = TRUE, main = "log(zn) ordinary kriging")
lzn.kriged.plot
# mapping the se
se.lzn.kriged.plot <- spplot(lzn.kriged["var1.var"], do.log = F, colorkey = TRUE, main ="var log(zn) ordinary kriging")
se.lzn.kriged.plot
```

### ordinary kriging mlr gstat
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# adding a column with log zinc
meuse <- meuse %>% dplyr::mutate(log_zinc = log(zinc))
# defining the regression task
task = makeRegrTask(id = "meuse",  data = meuse, target = "log_zinc")
task.krg <- dropFeatures(task = task, features = getTaskFeatureNames(task)[-c(1,2)])
# defining the learner
lrn.krg = makeLearner(cl = "regr.gstat", id = "ln(zn) mlr ordinary kriging", predict.type = "response", psill = 1, model = "Sph", range = 900, nugget = 1, locations = ~x+y)
# training the model
mod.krg = train(lrn.krg, task.krg)
# kriging
newdata.pred.krg = predict(object = mod.krg, newdata = meuse.grid)
mlr.krg <- bind_cols(data.frame(meuse.grid), newdata.pred.krg$data)
# mapping
coordinates(mlr.krg) <- ~x+y
gridded(mlr.krg) = TRUE
pred.plot <- spplot(mlr.krg["response"], do.log = T, colorkey = TRUE, main = mod.krg$learner$id)
pred.plot
# SE - defining the standard error learner by altering the previous one.
se.lrn.krg = setPredictType(lrn.krg, predict.type = "se")
# training the SE model
se.mod.krg = train(se.lrn.krg, task.krg)
# SE kriging
se.newdata.pred.krg = predict(object = se.mod.krg, newdata = meuse.grid)
se.mlr.krg <- bind_cols(data.frame(meuse.grid), se.newdata.pred.krg$data)
# SE mapping
coordinates(se.mlr.krg) <- ~x+y
gridded(se.mlr.krg) = TRUE
se.plot  <- spplot(se.mlr.krg["se"], do.log = T, colorkey = TRUE, main = se.mod.krg$learner$id)
se.plot 
```

### identical predictions ? 
```{r}
identical(lzn.kriged["var1.pred"]@data[[1]], mlr.krg["response"]@data[[1]])
```

## Kriging with External Drift (KED) = Universal Kriging (UK)
### raw gstat KED
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# making spatial
coordinates(meuse) = ~x+y
coordinates(meuse.grid) = ~x+y
gridded(meuse.grid) = TRUE
# computing sample variogram
lzn.vgm = variogram(log(zinc)~1, meuse)
# manually fitting a model to the vgm with a mean function where sqrt dist is the explanatory var
lzn.vgm = variogram(log(zinc)~sqrt(dist), meuse)
lzn.fit = fit.variogram(lzn.vgm, model = vgm(1, "Exp", 300, 1))
plot(lzn.vgm, lzn.fit)
# kriging
lzn.kriged = krige(log(zinc)~sqrt(dist), meuse, meuse.grid, model = lzn.fit)
# mapping
lzn.kriged.plot <- spplot(lzn.kriged["var1.pred"], do.log = F, colorkey = TRUE, main = "log(zn) kriging with external drift")
lzn.kriged.plot
# mapping the se
se.lzn.kriged.plot <- spplot(lzn.kriged["var1.var"], do.log = F, colorkey = TRUE, main = "var log(zn) kriging with external drift")
se.lzn.kriged.plot
```

### mlr gstat KED
```{r}
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# adding a column with log zinc
meuse <- meuse %>% dplyr::mutate(log_zinc = log(zinc))
# adding a column with sqrt dist
meuse <- meuse %>% dplyr::mutate(sqrt_dist = sqrt(dist))
meuse.grid <- meuse.grid %>% dplyr::mutate(sqrt_dist = sqrt(dist))
# defining the regression task
task = makeRegrTask(id = "meuse",  data = meuse, target = "log_zinc")
task.krg <- dropFeatures(task = task, features = getTaskFeatureNames(task)[-c(1,2,15)])
# defining the learner
lrn.krg = makeLearner(cl = "regr.gstat", id = "ln(zn) mlr kriging with external drift", predict.type = "response",  psill = 1, model = "Exp", range = 300, nugget = 1, locations = ~x+y) 
# training the model
mod.krg = train(lrn.krg, task.krg)
# kriging
newdata.pred.krg = predict(object = mod.krg, newdata = meuse.grid)
mlr.krg <- bind_cols(data.frame(meuse.grid), newdata.pred.krg$data)
# mapping
coordinates(mlr.krg) <- ~x+y
gridded(mlr.krg) = TRUE
pred.plot <- spplot(mlr.krg["response"], do.log = T, colorkey = TRUE, main = mod.krg$learner$id)
pred.plot
# SE - defining the standard error learner by altering the previous one.
se.lrn.krg = setPredictType(lrn.krg, predict.type = "se")
# training the SE model
se.mod.krg = train(se.lrn.krg, task.krg)
# SE kriging
se.newdata.pred.krg = predict(object = se.mod.krg, newdata = meuse.grid)
se.mlr.krg <- bind_cols(data.frame(meuse.grid), se.newdata.pred.krg$data)
# SE mapping
coordinates(se.mlr.krg) <- ~x+y
gridded(se.mlr.krg) = TRUE
se.plot  <- spplot(se.mlr.krg["se"], do.log = T, colorkey = TRUE, main = se.mod.krg$learner$id)
se.plot 
```

### identical predictions ? 
```{r}
identical(lzn.kriged["var1.pred"]@data[[1]], mlr.krg["response"]@data[[1]])
```

## interactively map prediction + SE with leaflet

Here we demonstrate how to map the prediction together with the uncertainty in an intuitive and intelligible way. Our approach is based on [this](http://spatial-analyst.net/wiki/index.php/Uncertainty_visualization#Visualization_of_uncertainty_using_whitening_in_R) post by *Tomislav Hengl* and [this](https://www.e-education.psu.edu/geog486/node/1891) post by *Penn state college of earth and mineral sciences*

```{r}
# For the example, using the output of mlr gstat KED from previous section
# loading the data
data(meuse)
data(meuse.grid)
# imputing values to missing data
meuse = impute(meuse, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
meuse.grid = impute(meuse.grid, classes = list(numeric = imputeMean(), factor = imputeMode()),
  dummy.classes = "integer")$data
# adding a column with log zinc
meuse <- meuse %>% dplyr::mutate(log_zinc = log(zinc))
# adding a column with sqrt dist
meuse <- meuse %>% dplyr::mutate(sqrt_dist = sqrt(dist))
meuse.grid <- meuse.grid %>% dplyr::mutate(sqrt_dist = sqrt(dist))
# defining the regression task
task = makeRegrTask(id = "meuse",  data = meuse, target = "log_zinc")
task.krg <- dropFeatures(task = task, features = getTaskFeatureNames(task)[-c(1,2,15)])
# defining the learner
lrn.krg = makeLearner(cl = "regr.gstat", id = "ln(zn) mlr kriging with external drift", predict.type = "response", psill = 1, model = "Exp", range = 300, nugget = 1, locations = ~x+y) 
# training the model
mod.krg = train(lrn.krg, task.krg)
# kriging
newdata.pred.krg = predict(object = mod.krg, newdata = meuse.grid)
mlr.krg <- bind_cols(data.frame(meuse.grid), newdata.pred.krg$data)
# mapping
coordinates(mlr.krg) <- ~x+y
gridded(mlr.krg) = TRUE
pred.plot <- spplot(mlr.krg["response"], do.log = T, colorkey = TRUE, main = mod.krg$learner$id)
pred.plot
# SE - defining the standard error learner by altering the previous one.
se.lrn.krg = setPredictType(lrn.krg, predict.type = "se")
# training the SE model
se.mod.krg = train(se.lrn.krg, task.krg)
# SE kriging
se.newdata.pred.krg = predict(object = se.mod.krg, newdata = meuse.grid)
se.mlr.krg <- bind_cols(data.frame(meuse.grid), se.newdata.pred.krg$data)

# mapping with leaflet

# loading the required libraries
library(leaflet)
library(dplyr)
library(htmltools)
library(sf)

# defining the function to create a palette of different levels of alpha for the choosen color 
alphaPal <- function(color) {
  alpha <- seq(0,1,0.1)
  r <- col2rgb(color, alpha=T)
  r <- t(apply(r, 1, rep, length(alpha)))
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  codes <- (rgb(r[1,], r[2,], r[3,], r[4,]))
  return(codes)
}

# keeping what we need 
interpolated.df <- se.mlr.krg %>% select(one_of(c("x","y","response","se")))
# making it spatial object class sf
interpolated.sf <- st_as_sf(interpolated.df,coords = c("x","y"))
# defining the crs
st_crs(interpolated.sf) <- 28992 # found at https://rdrr.io/cran/sp/man/meuse.html
# transforming to geographic CRS (EPSG = 4326)
interpolated.sf <- st_transform(interpolated.sf, crs = 4326)
# Now we need to inject this point info into polygon centered arounds the points to fake a raster layer but which is interactive
class(mlr.krg) # SpatialPixelsDataFrame
# making the gridded mlr.krg a raster 
grid.r <- raster::raster(mlr.krg, values=TRUE)
# convert raster to polygons
grid.sp = raster::rasterToPolygons(grid.r, dissolve = F)
class(grid.sp) # SpatialPolygonsDataFrame
# converting to sf for later joining
grid.sf <- st_as_sf(grid.sp)
st_crs(grid.sf) <- 28992 # found at https://rdrr.io/cran/sp/man/meuse.html
# transforming to geographic CRS (EPSG = 4326)
grid.sf <- st_transform(grid.sf, crs = 4326)
# injecting the prediction and se data into the polygon grid doing a spatial join
# interpolated.sf <- st_join(grid.sf, interpolated.sf) %>% select(one_of(c("response", "se")))
interpolated.pg.sf <- dplyr::bind_cols(grid.sf, interpolated.sf)
interpolated.pg.sf <- interpolated.pg.sf %>% select(one_of(c("response", "se")))
# Do we have polygons ? 
head(interpolated.pg.sf)

# defining the function to map using leaflet
leafletize <- function(data.sf){
  # to make the map responsive
  responsiveness.chr = "\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'"
  
  # defining the color palette for the response
  varPal <- colorNumeric(
    palette = "Spectral",
    domain = data.sf$response
  )
  
  # defining the transparent colorpal for the se
  uncPal <- colorNumeric(
    palette = alphaPal("#e6e6e6"),
    domain = data.sf$se,
    alpha = TRUE
  )

  # 
  prediction.map <- leaflet::leaflet(data.sf) %>%
    addProviderTiles(group = "Stamen",
                     providers$Stamen.Toner,
                     options = providerTileOptions(opacity = 0.25)
    ) %>%
    addProviderTiles(group = "Satellite",
                     providers$Esri.WorldImagery,
                     options = providerTileOptions(opacity = 1)
    ) %>%
    fitBounds(sf::st_bbox(data.sf)[[1]],
              sf::st_bbox(data.sf)[[2]],
              sf::st_bbox(data.sf)[[3]],
              sf::st_bbox(data.sf)[[4]]
    ) %>%
    addLayersControl(baseGroups = c("Stamen", "Satellite"),
                     overlayGroups = c("prediction", "se"),
                     options = layersControlOptions(collapsed = TRUE)
    ) %>%
    addEasyButton(easyButton(
      icon="fa-crosshairs", title="Locate Me",
      onClick=JS("function(btn, map){ map.locate({setView: true}); }"))) %>%
    htmlwidgets::onRender(paste0("
                                 function(el, x) {
                                 $('head').append(",responsiveness.chr,");
                                 }")
    ) %>%
    addPolygons(
                group = "prediction",
                color = "#444444", stroke = FALSE, weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 0.9,
                fillColor = ~varPal(response),
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE)
    )%>%
    addLegend(
              position = "bottomright", pal = varPal, values = ~response,
              title = "prediction",
              group = "prediction",
              opacity = 1
    )%>%
    addPolygons(
                group = "se",
                color = "#444444", stroke = FALSE, weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 1,
                fillColor = ~uncPal(se),
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE),
                label = ~ paste("prediction:", signif(data.sf$response, 2), "\n","se: ", signif(data.sf$se, 2))
    ) %>%
    addLegend(
              group = "se",
              position = "bottomleft", pal = uncPal, values = ~se,
              title = "se",
              opacity = 1
    )
  return(prediction.map)
}
# create the map object
prediction.map <- leafletize(interpolated.pg.sf)

# render it
html <- list(h3(paste0("interactive prediction map")),
                 prediction.map 
                 )
tagList(html)
```
