---
title: "Wrapper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message=FALSE}
library("mlr")
library("BBmisc")
library("ParamHelpers")
library("ggplot2")
library("lattice")

# show grouped code output instead of single lines
knitr::opts_chunk$set(collapse = TRUE)
```

Wrappers can be employed to extend integrated learners (`makeLearner()`)  with new functionality.
The broad scope of operations and methods which are implemented as wrappers underline the flexibility of the wrapping approach:

* [Data preprocessing](preproc.html){target="_blank"}
* [Imputation](impute.html){target="_blank"}
* [Bagging](bagging.html){target="_blank"}
* [Tuning](tune.html){target="_blank"}
* [Feature selection](feature_selection.html){target="_blank"}
* [Cost-sensitive classification](cost_sensitive_classif.html){target="_blank"}
* [Over- and undersampling](over_and_undersampling.html){target="_blank"} for imbalanced classification problems
* Multiclass extension (`makeMulticlassWrapper()`) for binary-class learners
* [Multilabel classification](multilabel.html){target="_blank"}

All these operations and methods have a few things in common:
First, they all wrap around `mlr` learners (`makeLearner()`)  and they return a new learner.
Therefore learners can be wrapped multiple times.
Second, they are implemented using a *train* (pre-model hook) and *predict* (post-model hook) method.

# Example: Bagging wrapper

In this section we exemplary describe the bagging wrapper to create a random forest which supports weights.
To achieve that we combine several decision trees from the `rpart` package to create our own custom random forest.

First, we create a weighted toy task.

```{r}
data(iris)
task = makeClassifTask(data = iris, target = "Species", weights = as.integer(iris$Species))
```

Next, we use `makeBaggingWrapper()` to create the base learners and the bagged learner.
We choose to set equivalents of `ntree` (100 base learners) and `mtry` (proportion of randomly selected features).

```{r}
base.lrn = makeLearner("classif.rpart")
wrapped.lrn = makeBaggingWrapper(base.lrn, bw.iters = 100, bw.feats = 0.5)
print(wrapped.lrn)
```

As we can see in the output, the wrapped learner inherited all properties from the base learner, especially the "weights" attribute is still present.
We can use this newly constructed learner like all base learners, i.e. we can use it in `train()`, `benchmark()`, `resample()`, etc.

```{r}
benchmark(tasks = task, learners = list(base.lrn, wrapped.lrn))
```

That far we are quite happy with our new learner.
But we hope for a better performance by tuning some hyperparameters of both the decision trees and bagging wrapper.
Let's have a look at the available hyperparameters of the fused learner:

```{r}
getParamSet(wrapped.lrn)
```

We choose to tune the parameters `minsplit` and `bw.feats` for the [mmce](measures.html){target="_blank"} using a random search (`TuneControl()`) in a 3-fold CV:

```{r}
ctrl = makeTuneControlRandom(maxit = 10)
rdesc = makeResampleDesc("CV", iters = 3)
par.set = makeParamSet(
  makeIntegerParam("minsplit", lower = 1, upper = 10),
  makeNumericParam("bw.feats", lower = 0.25, upper = 1)
)
tuned.lrn = makeTuneWrapper(wrapped.lrn, rdesc, mmce, par.set, ctrl)
print(tuned.lrn)
```

Calling the train method of the newly constructed learner performs the following steps:

1. The tuning wrapper sets parameters for the underlying model in slot `$next.learner` and calls its train method.
2. Next learner is the bagging wrapper. The passed down argument `bw.feats` is used in the bagging wrapper training function, the argument `minsplit` gets passed down to `$next.learner`. 
The base wrapper function calls the base learner `bw.iters` times and stores the resulting models.
3. The bagged models are evaluated using the mean [mmce](measures.html){target="_blank"} (default aggregation for this performance measure) and new parameters are selected using the tuning method.
4. This is repeated until the tuner terminates. Output is a tuned bagged learner.

```{r}
lrn = train(tuned.lrn, task = task)

print(lrn)
```

