Configure mlr
=============

If you really know what you are doing you might think *mlr* is limiting you.
Actually *mlr* is just designed to make errors due to misspellings or illogical parameter values as unlikely as possible.
But sometimes you want to brake those barriers to gain deeper access.
For all the parameters simply refer to the documentation [`?configureMlr`](http://berndbischl.github.io/mlr/man/configureMlr.html).

You are bothered by all the output on the console like in this example?
```{r}
library("mlr")

## Define a learning task and an appropriate learner
task = makeClassifTask(data = iris, target = "Species")
lrn = makeLearner("classif.ksvm")

## Perform a 3-fold cross-validation
rdesc = makeResampleDesc("CV", iters=3)
r = resample(lrn, task, rdesc)
```
Just try the following:
```{r}
configureMlr(show.learner.output=FALSE)
r = resample(lrn, task, rdesc)
```

One further scenario:
You want to access a new parameter of a learner which is already available through *mlr* but it is not implemented in *mlr* yet.
(If this is the case you might want to [contact us](https://github.com/berndbischl/mlr#get-in-touch) or [write an issue](https://github.com/berndbischl/mlr/issues/new) as well!)
But until then you can turn off *mlr*'s paramater checking like this:
```{r}
library("mlr")
lrn = makeLearner("classif.ksvm", newPar = 3)
lrn = makeLearner("classif.ksvm", epsilon = -3)
configureMlr(on.par.without.desc="quiet")
lrn = makeLearner("classif.ksvm", newPar = 3)
lrn = makeLearner("classif.ksvm", epsilon = -3)
```

