---
title: "Feature Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mlr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message=FALSE}
library("mlr")
library("BBmisc")
library("ParamHelpers")
library("ggplot2")
library("lattice")

# show grouped code output instead of single lines
knitr::opts_chunk$set(collapse = TRUE)
```

Often, data sets include a large number of features.
The technique of extracting a subset of relevant features is called feature selection.
Feature selection can enhance the interpretability of the model, speed up the learning process and improve the learner performance.
There exist different approaches to identify the relevant features.
`mlr` supports *filter* and *wrapper methods*.

# Filter methods

Filter methods assign an importance value to each feature.
Based on these values the features can be ranked and a feature subset can be selected.

## Calculating the feature importance

Different methods for calculating the feature importance are built into `mlr`'s function `generateFilterValuesData()` (`getFilterValues()` has been deprecated in favor of `generateFilterValuesData()`). 
Currently, classification, regression and survival analysis tasks are supported. 
A table showing all available methods can be found in article [filter methods](filter_methods.html){target="_blank"}.

Function `generateFilterValuesData()` requires the `Task()` and a character string specifying the filter method.

```{r}
fv = generateFilterValuesData(iris.task, method = "information.gain")
fv
```

``fv`` is a `FilterValues()` object and ``fv$data`` contains a `data.frame` that gives the importance values for all features. 
Optionally, a vector of filter methods can be passed.

```{r}
fv2 = generateFilterValuesData(iris.task, method = c("information.gain", "chi.squared"))
fv2$data
```

A bar plot of importance values for the individual features can be obtained using function `plotFilterValues()`.

```{r}
plotFilterValues(fv2)
```

By default `plotFilterValues()` will create facetted subplots if multiple filter methods are passed as input to `generateFilterValuesData()`.

There is also an experimental `ggvis` plotting function, `plotFilterValuesGGVIS()`. 
This takes the same arguments as `plotFilterValues()` and produces a `shiny` application that allows the interactive selection of the displayed filter method, the number of features selected, and the sorting method (e.g., ascending or descending).

```{r, eval=FALSE}
plotFilterValuesGGVIS(fv2)
```

According to the ``"information.gain"`` measure, `Petal.Width` and `Petal.Length` contain the most information about the target variable `Species`.

## Selecting a feature subset

With `mlr`'s function `filterFeatures()` you can create a new `Task()` by leaving out features of lower importance.

There are several ways to select a feature subset based on feature importance values:

* Keep a certain *absolute number* (``abs``) of features with highest importance.
* Keep a certain *percentage* (``perc``) of features with highest importance.
* Keep all features whose importance exceeds a certain *threshold value* (``threshold``).

Function `filterFeatures()` supports these three methods as shown in the following example.
Moreover, you can either specify the ``method`` for calculating the feature importance or you can use previously computed importance values via argument ``fval``.

```{r}
# Keep the 2 most important features
filtered.task = filterFeatures(iris.task, method = "information.gain", abs = 2)

# Keep the 25% most important features
filtered.task = filterFeatures(iris.task, fval = fv, perc = 0.25)

# Keep all features with importance greater than 0.5
filtered.task = filterFeatures(iris.task, fval = fv, threshold = 0.5)
filtered.task
```

## Fuse a learner with a filter method

Often feature selection based on a filter method is part of the data preprocessing and in a subsequent step a learning method is applied to the filtered data.
In a proper experimental setup you might want to automate the selection of the features so that it can be part of the validation method of your choice.
A Learner (`makeLearner()`) can be fused with a filter method by function `makeFilterWrapper()`.
The resulting Learner (`makeLearner()`) has the additional class attribute `FilterWrapper()`.

In the following example we calculate the 10-fold cross-validated error rate [mmce](measures.html){target="_blank"} of the k-nearest neighbor classifier (`FNN::fnn()`) with preceding feature selection on the `iris` (`datasets::iris()`) data set.
We use ``"information.gain"`` as importance measure and select the 2 features with highest importance.
In each resampling iteration feature selection is carried out on the corresponding training data set before fitting the learner.

```{r}
lrn = makeFilterWrapper(learner = "classif.fnn", fw.method = "information.gain", fw.abs = 2)
rdesc = makeResampleDesc("CV", iters = 10)
r = resample(learner = lrn, task = iris.task, resampling = rdesc, show.info = FALSE, models = TRUE)
r$aggr
```

You may want to know which features have been used. Luckily, we have called
`resample()` with the argument `models = TRUE`, which means that `r$models`
contains a `list` of models (`makeWrappedModel()`) fitted in the individual resampling iterations.
In order to access the selected feature subsets we can call `getFilteredFeatures()` on each model.

```{r}
sfeats = sapply(r$models, getFilteredFeatures)
table(sfeats)
```

The selection of features seems to be very stable.
The features `Sepal.Length` and `Sepal.Width` did not make it into a single fold.

## Tuning the size of the feature subset

In the above examples the number/percentage of features to select or the threshold value have been arbitrarily chosen.
If filtering is a preprocessing step before applying a learning method optimal values with regard to the learner performance can be found by [tuning](tune.html){target="_blank"}.

In the following regression example we consider the `BostonHousing` (`mlbench::BostonHousing()`) data set.
We use a linear regression model and determine the optimal percentage value for feature selection such that the 3-fold cross-validated mean squared error (`mse()`) of the learner is minimal.
As search strategy for tuning a grid search is used.

```{r}
lrn = makeFilterWrapper(learner = "regr.lm", fw.method = "chi.squared")
ps = makeParamSet(makeDiscreteParam("fw.perc", values = seq(0.2, 0.5, 0.05)))
rdesc = makeResampleDesc("CV", iters = 3)
res = tuneParams(lrn, task = bh.task, resampling = rdesc, par.set = ps,
  control = makeTuneControlGrid())
res
```

The performance of all percentage values visited during tuning is:

```{r}
as.data.frame(res$opt.path)
```

The optimal percentage and the corresponding performance can be accessed as follows:

```{r}
res$x
res$y
```

After tuning we can generate a new wrapped learner with the optimal percentage value for further use.

```{r}
lrn = makeFilterWrapper(learner = "regr.lm", fw.method = "chi.squared", fw.perc = res$x$fw.perc)
mod = train(lrn, bh.task)
mod

getFilteredFeatures(mod)
```

Here is another example using multi-criteria [tuning](tune.html){target="_blank"}.
We consider linear discriminant analysis (`MASS::lda()`) with precedent feature selection based on the Chi-squared statistic of independence (``"chi.squared"``) on the `Sonar` (`mlbench::sonar()`) data set and tune the threshold value.
During tuning both, the false positive and the false negative rate [fpr](measures.html){target="_blank"} and [fnr](measures.html){target="_blank"}), are minimized. 
As search strategy we choose a random search (see `makeTuneMultiCritControlRandom` (`?TuneMultiCritControl()`).

```{r}
lrn = makeFilterWrapper(learner = "classif.lda", fw.method = "chi.squared")
ps = makeParamSet(makeNumericParam("fw.threshold", lower = 0.1, upper = 0.9))
rdesc = makeResampleDesc("CV", iters = 10)
res = tuneParamsMultiCrit(lrn, task = sonar.task, resampling = rdesc, par.set = ps,
  measures = list(fpr, fnr), control = makeTuneMultiCritControlRandom(maxit = 50L),
  show.info = FALSE)
res
head(as.data.frame(res$opt.path))
```

The results can be visualized with function `plotTuneMultiCritResult()`.
The plot shows the false positive and false negative rates for all parameter values visited during tuning. 
The size of the points on the Pareto front is slightly increased.

```{r}
plotTuneMultiCritResult(res)
```

# Wrapper methods

Wrapper methods use the performance of a learning algorithm to assess the usefulness of a feature set.
In order to select a feature subset a learner is trained repeatedly on different feature subsets and the subset which leads to the best learner performance is chosen.

In order to use the wrapper approach we have to decide:

* How to assess the performance: This involves choosing a performance measure that serves as feature selection criterion and a resampling strategy.
* Which learning method to use.
* How to search the space of possible feature subsets.

The search strategy is defined by functions following the naming convention
``makeFeatSelControl<search_strategy``.
The following search strategies are available:

* Exhaustive search `makeFeatSelControlExhaustive` (`?FeatSelControl()`),
* Genetic algorithm `makeFeatSelControlGA` (`?FeatSelControl()`),
* Random search `makeFeatSelControlRandom` (`?FeatSelControl()`),
* Deterministic forward or backward search `makeFeatSelControlSequential` (`?FeatSelControl()`).

## Select a feature subset

Feature selection can be conducted with function `selectFeatures()`.

In the following example we perform an exhaustive search on the
`Wisconsin Prognostic Breast Cancer` (`TH.data::wpbc()`) data set.
As learning method we use the `Cox proportional hazards model` (`survival::coxph()`).
The performance is assessed by the holdout estimate of the concordance index [cindex](measures.html){target="_blank"}).

```{r}
# Specify the search strategy
ctrl = makeFeatSelControlRandom(maxit = 20L)
ctrl
```

``ctrl`` is a`FeatSelControl()` object that contains information about the search strategy and potential parameter values.

```{r}
# Resample description
rdesc = makeResampleDesc("Holdout")

# Select features
sfeats = selectFeatures(learner = "surv.coxph", task = wpbc.task, resampling = rdesc,
  control = ctrl, show.info = FALSE)
sfeats
```

``sfeats``is a `FeatSelResult` (`selectFeatures()`) object.
The selected features and the corresponding performance can be accessed as follows:

```{r}
sfeats$x
sfeats$y
```


In a second example we fit a simple linear regression model to the `BostonHousing` (`mlbench::BostonHousing()`) data set and use a sequential search to find a feature set that minimizes the mean squared error [mse](measures.html){target="_blank"}).
``method = "sfs"`` indicates that we want to conduct a sequential forward search where features are added to the model until the performance cannot be improved anymore.
See the documentation page `makeFeatSelControlSequential` (`?FeatSelControl()`) for other available sequential search methods.
The search is stopped if the improvement is smaller than ``alpha = 0.02``.

```{r}
# Specify the search strategy
ctrl = makeFeatSelControlSequential(method = "sfs", alpha = 0.02)

# Select features
rdesc = makeResampleDesc("CV", iters = 10)
sfeats = selectFeatures(learner = "regr.lm", task = bh.task, resampling = rdesc, control = ctrl,
  show.info = FALSE)
sfeats
```

Further information about the sequential feature selection process can be obtained by function `analyzeFeatSelResult()`.

```{r}
analyzeFeatSelResult(sfeats)
```

## Fuse a learner with feature selection

A Learner (`makeLearner()`) can be fused with a feature selection strategy (i.e., a search strategy, a performance measure and a resampling strategy) by function `makeFeatSelWrapper()`.
During training features are selected according to the specified selection scheme. 
Then, the learner is trained on the selected feature subset.

```{r}
rdesc = makeResampleDesc("CV", iters = 3)
lrn = makeFeatSelWrapper("surv.coxph", resampling = rdesc,
  control = makeFeatSelControlRandom(maxit = 10), show.info = FALSE)
mod = train(lrn, task = wpbc.task)
mod
```

The result of the feature selection can be extracted by function `getFeatSelResult()`.

```{r}
sfeats = getFeatSelResult(mod)
sfeats
```

The selected features are:

```{r}
sfeats$x
```

The 5-fold cross-validated performance of the learner specified above can be computed as follows:

```{r}
out.rdesc = makeResampleDesc("CV", iters = 5)

r = resample(learner = lrn, task = wpbc.task, resampling = out.rdesc, models = TRUE,
  show.info = FALSE)
r$aggr
```

The selected feature sets in the individual resampling iterations can be extracted as follows:

```{r}
lapply(r$models, getFeatSelResult)
```
